---
title: "Weather for CIMMYT experimental station, Harare, Zimbabwe"
author:
- João Vasco Silva \& Mainassara Zaman-Allah
- International Maize and Wheat Improvement Center (CIMMYT)
date: "`r format(Sys.time(), '%d-%B-%Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: yes
    fig_caption: yes
    css: style.css
    header-includes:
    - \usepackage{setspace}
    - \sinlgespacing
---

<body style="background-color:rgb(249, 247, 245);">

```{r, echo=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, echo=FALSE, warning=FALSE}
# package names
packages <- c("leaflet", "plotly", "highcharter", "visNetwork", "DT", "dygraphs", "xts", "htmltools", "tidyverse", "manipulateWidget", "downloadthis")

# install packages
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == FALSE)){
  install.packages(packages[!installed_packages])}

# load packages
invisible(lapply(packages, library, character.only = TRUE))
```

<hr style="border:1.5px solid black">
# **Background information**

This notebook summarizes time series weather data, starting in 2009 until 2019, from an automated weather station located in CIMMYT Southern Africa Regional Office in Harare, Zimbabwe. The weather station contains sensors to measure solar radiation, minimum and maximum temperature, relative humidity, and precipitation (among others). Reference evapotranspiration was estimated for the location using the Penman-Monteith equations. The weather station is maintained by dr Mainassara Zaman-Allah (please contact Z.MainassaraAbdou@cgiar.org for further details).

The weather data available through this weather station is currently being used to (1) inform the maize breeding program of CIMMYT in Southern Africa, and (2) monitor weather variability and extremes in agronomy trials. The same data could be further used to simulate potential and water-limited yields with crop growth models, among other applications in plant breeding and agronomy. This notebook is updated yearly as data become available.

<hr style="border:1.5px solid black">
# **Location of weather station**

The weather station is installed in CIMMYT Southern African Regional Office, a highland area with high potential for agricultural production, in the outskirts North of Harare (GPS coordinates: -17.72 S, 31.02 E, 1490 m above sea level). The CIMMYT experimental station is part of the research farm from the University of Zimbabwe. The main crops cultivated in the surrounding farms include maize, wheat, potato, and a range of horticultural crops.

```{r, echo=FALSE, fig.align="center", fig.height=3.25, out.width="100%"}
location <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')
leaflet(location) %>% 
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addCircleMarkers()
```

The region is characterized by a Dry-winter Subtropical Highland Climate (Cwb), according the Köppen climate classification (https://en.wikipedia.org/wiki/K%C3%B6ppen_climate_classification). The Dry-winter Subtropical Highland Climate is found in the highlands within the tropics of Central America, South America, Africa, and Asia. It exists mostly within a band within 20° of the equator, in both hemispheres, though it can sometimes be found out to 40°. The summers are mild and very rainy with a monthly average temperature below 22 °C but above freezing. At least one month has an average temperature below 18 °C. The winters are dry and noticeably different from summers. However, they are more like spring than a traditional winter. Despite its low latitude, higher altitudes allow these regions to share climate characteristics with oceanic climates. If it were not for this elevation difference, these areas would be classified as either tropical or humid subtropical climates.Though this climate exists in tropical regions, it may too cool for some tropical plants to grow. Temperate crops may be more suitable. 

<hr style="border:1.5px solid black">
# **Daily weather conditions**

```{r, echo=FALSE, fig.height=10, fig.asp=0.618, out.width="100%", fig.align="center"}

# plot 1
data <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')
data_aggr <- aggregate(data[c('IRRAD_MJm2day', 'RELHUM_mean_perc')], by=list('year'=data$Year, 'month'=data$Month), FUN=mean, na.rm=T)
ts_data_rad <- xts(data[,15], order.by=as.Date(data[,8], "%Y-%m-%d"))
ts_data_hum <- xts(data[,28], order.by=as.Date(data[,8], "%Y-%m-%d"))
ts_rad <- cbind(ts_data_rad, ts_data_hum)
rad <- dygraph(ts_rad, group=ts_rad) %>%
  dyAxis("y", label = "Solar radiation (MJ/m2/day)", valueRange = c(0, 55), independentTicks=TRUE, axisLabelWidth=35)%>%
  dyAxis("y2", label = "Relative humidity (%)", valueRange = c(0, 110), independentTicks=TRUE, axisLabelWidth=35) %>%
  dySeries("ts_data_rad", label='Radiation', color='goldenrod', axis='y') %>%
  dySeries("ts_data_hum", label='Humidity', color='steelblue', axis='y2') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2) %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE)

# plot 2
data <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')
data_aggr <- aggregate(data[c('TMIN', 'TMAX')], by=list('year'=data$Year, 'month'=data$Month), FUN=mean, na.rm=T)
ts_data_tmin <- xts(data[,17], order.by=as.Date(data[,8], "%Y-%m-%d"))
ts_data_tmax <- xts(data[,16], order.by=as.Date(data[,8], "%Y-%m-%d"))
ts_temp <- cbind(ts_data_tmin, ts_data_tmax)
temp <- dygraph(ts_temp, group=ts_rad) %>%
  dyAxis("y", label='Air temperature (C)', valueRange = c(-5, 45), axisLabelWidth=35) %>% 
  dyAxis("y2", label='Air temperature (C)', valueRange = c(-5, 45), axisLabelWidth=35) %>% 
  dySeries("ts_data_tmax", label="TMAX", axis='y') %>%
  dySeries("ts_data_tmin", label="TMIN", axis='y2') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2) %>%
  dyLimit(35, "", strokePattern="solid", color="red") %>% 
  dyLimit(0, "", strokePattern="solid", color="red") %>% 
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE)

# plot 3
data <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')
data_aggr <- aggregate(data[c('RAIN_mm', 'ET0_mm')], by=list('year'=data$Year, 'month'=data$Month), FUN=mean, na.rm=T)
ts_data_rain <- xts(data[,21], order.by=as.Date(data[,8], "%Y-%m-%d"))
ts_data_et0 <- xts(data[,23], order.by=as.Date(data[,8], "%Y-%m-%d"))
ts_rain <- cbind(ts_data_rain, ts_data_et0)
rain <- dygraph(ts_rain, group=ts_rad) %>%
  dyAxis("y", label = "Precipitation (mm)", valueRange = c(0, 60), independentTicks = TRUE, axisLabelWidth=35)%>%
  dyAxis("y2", label = "Reference ET (mm)", valueRange = c(0, 12), independentTicks = FALSE, axisLabelWidth=35) %>%
  dySeries("ts_data_rain", label='RAIN', color='darkblue', axis='y') %>%
  dySeries("ts_data_et0", label='ET0', color='dodgerblue', axis='y2') %>%  
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2) %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyRangeSelector(height = 40, dateWindow = c("2018-01-01", "2018-12-31"))

# both
combineWidgets(rad, temp, rain, nrow=3, ncol=1, rowsize=c(1,1,1), height=720)
```

<hr style="border:1.5px solid black">
# **Climate change over time**

```{r, echo=FALSE, fig.height=10, fig.asp=0.3, out.width="100%", fig.align="center"}

# radiation & rain -------------------------------------------------------------

# load data
data <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')

# january - december
data_yr <- data
data_yr <- aggregate(data_yr[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_yr$Year), FUN=sum, na.rm=T)
data_yr$year <- paste0(data_yr$year, '-01-01')
data_yr_rain_cum <- xts(data_yr[,2], order.by=as.Date(data_yr$year))  
data_yr_srad_cum <- xts(data_yr[,3], order.by=as.Date(data_yr$year))  

# january - march
data_mar <- subset(data, Month == 1 | Month == 2 | Month == 3)
data_mar <- aggregate(data_mar[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_mar$Year), FUN=sum, na.rm=T)
data_mar$year <- paste0(data_mar$year, '-01-01')
data_mar_rain_cum <- xts(data_mar[,2], order.by=as.Date(data_mar$year))  
data_mar_srad_cum <- xts(data_mar[,3], order.by=as.Date(data_mar$year))  

# january - june
data_jun <- subset(data, Month == 1 | Month == 2 | Month == 3 | Month == 4 | Month == 5 | Month == 6)
data_jun <- aggregate(data_jun[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_jun$Year), FUN=sum, na.rm=T)
data_jun$year <- paste0(data_jun$year, '-01-01')
data_jun_rain_cum <- xts(data_jun[,2], order.by=as.Date(data_jun$year))  
data_jun_srad_cum <- xts(data_jun[,3], order.by=as.Date(data_jun$year))  

# january - september
data_sep <- subset(data, Month == 1 | Month == 2 | Month == 3 | Month == 4 | Month == 5 | Month == 6 | Month == 7 | Month == 8 | Month == 9)
data_sep <- aggregate(data_sep[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_sep$Year), FUN=sum, na.rm=T)
data_sep$year <- paste0(data_sep$year, '-01-01')
data_sep_rain_cum <- xts(data_sep[,2], order.by=as.Date(data_sep$year))  
data_sep_srad_cum <- xts(data_sep[,3], order.by=as.Date(data_sep$year))  

# data for plot
data_plot_srad <- cbind(data_mar_srad_cum/100, data_jun_srad_cum/100, data_sep_srad_cum/100, data_yr_srad_cum/100)
p1 <- dygraph(data_plot_srad) %>%
  dyAxis("y", label = "Solar radiation (x100 MJ/m2)", valueRange = c(0, 100), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_srad_cum.100", label='Year', color='orangered', axis='y') %>%
  dySeries("data_mar_srad_cum.100", label='Jan-Mar', color='goldenrod', axis='y') %>%
  dySeries("data_jun_srad_cum.100", label='Apr-Jun', color='goldenrod', axis='y') %>%
  dySeries("data_sep_srad_cum.100", label='Jul-Sep', color='goldenrod', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  # dyRangeSelector(height = 40, dateWindow = c("2009-01-01", "2021-01-01")) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# data for plot
data_plot_rain <- cbind(data_mar_rain_cum/10, data_jun_rain_cum/10, data_sep_rain_cum/10, data_yr_rain_cum/10)
p2 <- dygraph(data_plot_rain, group=data_plot_srad) %>%
  dyAxis("y", label = "Precipitation (cm)", valueRange = c(0, 125), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_rain_cum.10", label='Year', color='darkblue', axis='y') %>%
  dySeries("data_mar_rain_cum.10", label='Jan-Mar', color='dodgerblue', axis='y') %>%
  dySeries("data_jun_rain_cum.10", label='Apr-Jun', color='dodgerblue', axis='y') %>%
  dySeries("data_sep_rain_cum.10", label='Jul-Sep', color='dodgerblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  # dyRangeSelector(height = 40, dateWindow = c("2009-01-01", "2021-01-01")) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# temperature ------------------------------------------------------------------

# january - december
data_yr <- data
data_yr <- aggregate(data_yr[c('TMAX', 'TMIN')], by=list('year'=data_yr$Year), FUN=mean, na.rm=T)
data_yr$year <- paste0(data_yr$year, '-01-01')
data_yr_tmax_av <- xts(data_yr[,2], order.by=as.Date(data_yr$year))  
data_yr_tmin_av <- xts(data_yr[,3], order.by=as.Date(data_yr$year))  

# january - march
data_mar <- subset(data, Month == 1 | Month == 2 | Month == 3)
data_mar <- aggregate(data_mar[c('TMAX', 'TMIN')], by=list('year'=data_mar$Year), FUN=mean, na.rm=T)
data_mar$year <- paste0(data_mar$year, '-01-01')
data_mar_tmax_av <- xts(data_mar[,2], order.by=as.Date(data_mar$year))  
data_mar_tmin_av <- xts(data_mar[,3], order.by=as.Date(data_mar$year))  

# april - june
data_jun <- subset(data, Month == 4 | Month == 5 | Month == 6)
data_jun <- aggregate(data_jun[c('TMAX', 'TMIN')], by=list('year'=data_jun$Year), FUN=mean, na.rm=T)
data_jun$year <- paste0(data_jun$year, '-01-01')
data_jun_tmax_av <- xts(data_jun[,2], order.by=as.Date(data_jun$year))  
data_jun_tmin_av <- xts(data_jun[,3], order.by=as.Date(data_jun$year))  

# july - september
data_sep <- subset(data, Month == 7 | Month == 8 | Month == 9)
data_sep <- aggregate(data_sep[c('TMAX', 'TMIN')], by=list('year'=data_sep$Year), FUN=mean, na.rm=T)
data_sep$year <- paste0(data_sep$year, '-01-01')
data_sep_tmax_av <- xts(data_sep[,2], order.by=as.Date(data_sep$year))  
data_sep_tmin_av <- xts(data_sep[,3], order.by=as.Date(data_sep$year))  

# october - december
data_dec <- subset(data, Month == 10 | Month == 11 | Month == 12)
data_dec <- aggregate(data_dec[c('TMAX', 'TMIN')], by=list('year'=data_dec$Year), FUN=mean, na.rm=T)
data_dec$year <- paste0(data_dec$year, '-01-01')
data_dec_tmax_av <- xts(data_dec[,2], order.by=as.Date(data_dec$year))  
data_dec_tmin_av <- xts(data_dec[,3], order.by=as.Date(data_dec$year))  

# data for plot
data_plot_tmax <- cbind(data_mar_tmax_av, data_jun_tmax_av, data_sep_tmax_av, data_dec_tmax_av, data_yr_tmax_av)
p3 <- dygraph(data_plot_tmax) %>%
  dyAxis("y", label = "Maximum temperature (C)", valueRange = c(15, 35), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_tmax_av", label='Year', color='black', axis='y') %>%
  dySeries("data_mar_tmax_av", label='Jan-Mar', color='red', axis='y') %>%
  dySeries("data_jun_tmax_av", label='Apr-Jun', color='salmon', axis='y') %>%
  dySeries("data_sep_tmax_av", label='Jul-Sep', color='plum', axis='y') %>%
  dySeries("data_dec_tmax_av", label='Oct-Dec', color='firebrick', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  # dyRangeSelector(height = 40, dateWindow = c("2009-01-01", "2021-01-01")) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# data for plot
data_plot_tmin <- cbind(data_mar_tmin_av, data_jun_tmin_av, data_sep_tmin_av, data_dec_tmin_av, data_yr_tmin_av)
p4 <- dygraph(data_plot_tmin) %>%
  dyAxis("y", label = "Minimum temperature (C)", valueRange = c(5, 25), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_tmin_av", label='Year', color='black', axis='y') %>%
  dySeries("data_mar_tmin_av", label='Jan-Mar', color='blue', axis='y') %>%
  dySeries("data_jun_tmin_av", label='Apr-Jun', color='steelblue', axis='y') %>%
  dySeries("data_sep_tmin_av", label='Jul-Sep', color='skyblue', axis='y') %>%
  dySeries("data_dec_tmin_av", label='Oct-Dec', color='royalblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  # dyRangeSelector(height = 40, dateWindow = c("2009-01-01", "2021-01-01")) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# number of days ---------------------------------------------------------------

data <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')
heat_days <- subset(data, TMAX > 30)
heat_days$unique <- 1
heat_days <- aggregate(heat_days[c('unique')], by=list('year'=heat_days$Year), FUN=sum, na.rm=T)
heat_days$year <- paste0(heat_days$year, '-01-01')
heat_days <- xts(heat_days[,2], order.by=as.Date(heat_days$year))  
cold_days <- subset(data, TMIN < 10)
cold_days$unique <- 1
cold_days <- aggregate(cold_days[c('unique')], by=list('year'=cold_days$Year), FUN=sum, na.rm=T)
cold_days$year <- paste0(cold_days$year, '-01-01')
cold_days <- xts(cold_days[,2], order.by=as.Date(cold_days$year))  

# data for plot
data_plot_days <- cbind(heat_days, cold_days)
p5 <- dygraph(data_plot_days) %>%
  dyAxis("y", label = "Number of heat & cold days", valueRange = c(0, 120), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("heat_days", label='Heat days (TMAX > 30)', color='firebrick', axis='y') %>%
  dySeries("cold_days", label='Cold days (TMIN < 10)', color='royalblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  # dyRangeSelector(height = 40, dateWindow = c("2009-01-01", "2021-01-01")) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# number of days ---------------------------------------------------------------

rain_days <- subset(data, RAIN_mm > 0)
rain_days$unique <- 1
rain_days <- aggregate(rain_days[c('unique')], by=list('year'=rain_days$Year), FUN=sum, na.rm=T)
rain_days$year <- paste0(rain_days$year, '-01-01')
rain_days <- xts(rain_days[,2], order.by=as.Date(rain_days$year))  

# data for plot
data_plot_days <- cbind(heat_days, cold_days, rain_days)[,c(3)]
p6 <- dygraph(data_plot_days) %>%
  dyAxis("y", label = "Number of rain days", valueRange = c(60, 160), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("rain_days", label='Rain days (Rain > 0)', color='royalblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  # dyRangeSelector(height = 40, dateWindow = c("2009-01-01", "2021-01-01")) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }")

# all plots --------------------------------------------------------------------

# both
combineWidgets(p1, p2, p3, p4, p5, p6, nrow=3, ncol=2, rowsize=c(1), height=550+275)

#plotInteractive <- function(Months, Variable){
#  # subset
#  month_subset <- subset(data, Month %in% Months)
#  data_yr <- aggregate(month_subset[c(Variable)], by=list('year'=month_subset$Year), FUN=mean, na.rm=T)
#  # plot
#  dygraph(data_yr, main = Variable) %>%
#    dyAxis("y", label = Variable) %>% #, valueRange = c(12, 18)) %>%
#    dySeries(Variable, label=Variable, axis='y') %>%
#    dyOptions(axisLineWidth = 1.5, fillGraph = TRUE, drawGrid = TRUE, drawPoints=T, pointSize=2) %>%
#    dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) }
# manipulateWidget(
#  plotInteractive(Months, Variable),
#  Variable = mwSelect(sort(unique(c('IRRAD_MJm2day', 'RAIN_mm'))), "IRRAD_MJm2day"),
#  Months = mwCheckboxGroup(unique(data$Month), 0)
# )

```

<hr style="border:1.5px solid black">
# **Cumulative values**

```{r, echo=FALSE, fig.height=10, fig.asp=0.3, out.width="100%", fig.align="center"}

data <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')

# SRAD
srad <- data[c('Y_year', 'DOY', 'IRRAD_MJm2day')]
srad_final <- c()
for(yr in unique(srad$Y_year)){
  srad_yr <- subset(srad, Y_year == yr)
  srad_yr$IRRAD_MJm2day <- cumsum(srad_yr$IRRAD_MJm2day)
  srad_final <- rbind(srad_final, srad_yr)}
srad_final$IRRAD_MJm2day <- srad_final$IRRAD_MJm2day/100
srad <- reshape2::dcast(srad_final, DOY ~ Y_year, value.var='IRRAD_MJm2day')
cum_srad <- dygraph(srad) %>%
  dyGroup(c(names(srad)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::inferno(14)) %>%
  dyAxis("y", label = "Solar radiation (x100 MJ/m2/d)", valueRange = c(0, 100), independentTicks = TRUE, axisLabelWidth=35) %>%
  dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyLegend(width=300)

# GDD
gdd <- data[c('Y_year', 'DOY', 'GDD')]
gdd_final <- c()
for(yr in unique(gdd$Y_year)){
  gdd_yr <- subset(gdd, Y_year == yr)
  gdd_yr$GDD <- cumsum(gdd_yr$GDD)
  gdd_final <- rbind(gdd_final, gdd_yr)}
gdd_final$GDD <- gdd_final$GDD/100
gdd <- reshape2::dcast(gdd_final, DOY ~ Y_year, value.var='GDD')
gdd_val <- dygraph(gdd) %>%
  dyGroup(c(names(gdd)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::inferno(14)) %>%
  dyAxis("y", label = "Growing degrees day (x100)", valueRange = c(0, 95), independentTicks = TRUE, axisLabelWidth=35) %>%
  dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyLegend(width=300)

# RAIN
rain <- data[c('Y_year', 'DOY', 'RAIN_mm')]
rain_final <- c()
for(yr in unique(rain$Y_year)){
  rain_yr <- subset(rain, Y_year == yr)
  rain_yr$RAIN_mm <- cumsum(rain_yr$RAIN_mm)
  rain_final <- rbind(rain_final, rain_yr)}
rain_final$RAIN_mm <- rain_final$RAIN_mm/10
rain <- reshape2::dcast(rain_final, DOY ~ Y_year, value.var='RAIN_mm')
cum_rain <- dygraph(rain) %>%
  dyGroup(c(names(rain)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::viridis(14)) %>%
  dyAxis("y", label = "Precipitation (cm)", valueRange = c(0, 130), independentTicks = TRUE, axisLabelWidth=35) %>%
  dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyLegend(width=300)

# ET
et <- data[c('Y_year', 'DOY', 'ET0_mm')]
et_final <- c()
for(yr in unique(et$Y_year)){
  et_yr <- subset(et, Y_year == yr)
  et_yr$ET0_mm <- cumsum(et_yr$ET0_mm)
  et_final <- rbind(et_final, et_yr)}
et_final$ET0_mm <- et_final$ET0_mm/10
et <- reshape2::dcast(et_final, DOY ~ Y_year, value.var='ET0_mm')
cum_et <- dygraph(et) %>%
   dyGroup(c(names(et)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::viridis(14)) %>%
   dyAxis("y", label = "Reference ET (cm)", valueRange = c(0, 320), independentTicks = TRUE, axisLabelWidth=35) %>%
   dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
   dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
   dyAxis("x", drawGrid = FALSE) %>%
   dyLegend(width=300)

# pal <- colorRampPalette(c('darkred', "darkorange", 'orange', 'yellowgreen', 'forestgreen', 'darkgreen'))
# cum_et <- hchart(et_final, "line", hcaes(x=DOY, y=ET0_mm, group=Year, color=Year)) %>%
#   hc_colors(c(pal(length(unique(et_final$Year))))) %>%
#   hc_legend() %>%
#   hc_yAxis(title = list(text="Reference ET (mm)", style=list(color='black', fontSize=16)),
#            labels = list(style=list(color='black', fontSize=15))) %>%  
#   hc_xAxis(title = list(text="Day-of-the-year (DOY)", style=list(color='black', fontSize=16)), 
#            labels = list(style=list(color='black', fontSize=15)),  
#            tickInterval = 1) %>%
#   hc_chart(events = list(load = JS("function() {
#       var chart = this; 
#       chart.update({ chart: { plotBorderColor: 'black', plotBorderWidth: '1' } } ); }")))

# both
combineWidgets(cum_rain, gdd_val, cum_et, cum_srad, nrow=2, ncol=2, rowsize=c(1), height=600+100) # cum_et

```

<hr style="border:1.5px solid black">
# **Table with complete data**

```{r, echo=FALSE, warning=FALSE}
data <- read.csv('D:/# Jvasco/Working Papers/CIMMYT Yield Potential/3-MAIZE-Zimbabwe/Maina/CIMMYT SARO/2-rmarkdown/weather-data-harare.csv')
data <- data[c('Station', 'date', 'Day', 'Month', 'Year', 'DOY', 'IRRAD_MJm2day', 
               'RELHUM_mean_perc', 'TMAX', 'TMIN', 'GDD', 'RAIN_mm', 'ET0_mm',
               'WIND_ms', 'VAP_kPa', 'LAT', 'LONG', 'Contact')]
colnames(data)[7] <- 'SRAD'
colnames(data)[8] <- 'RELHUM'
round_df <- function(x, digits) {
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x}
data <- round_df(data, 2)
data %>%
   DT::datatable(options=list(scrollX=TRUE, scrollCollapse=TRUE), rownames = FALSE)
```

```{css, echo=FALSE, warning=FALSE}
.button_large {font-size: 18px;}
```

```{r, echo=FALSE, warning=FALSE}
# data %>%
#  download_this(
#    output_name = "Weather_data_cimmyt_harare",
#    output_extension = ".xlsx",
#    button_label = "DOWNLOAD DATA AS MS-EXCEL",
#    button_type = "default",
#    has_icon = TRUE,
#    icon = "fa fa-save",
#    class = "button_large")
```

```{r, echo=FALSE, warning=FALSE}
htmltools::img(src = knitr::image_uri("CIMMYT-logo.png"), 
               alt = 'logo', 
               style = 'float: right;padding-right:10px;width:450px')
```
